\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-02-14 analysis/manuscript.Rmd
continue introduction ii

For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
The idea was that when journals publish more and more open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access. 
However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
Other publishers also recorded a low uptake.
In 2011, only 1-2% of articles in over 4,300 hybrid journals from large publishers were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].

With the introduction of central funding mechanisms for hybrid open access publication fees in a few countries since 2012, notably in the UK, an increase in open access could be observed [@Bj_rk_2017; @Jubb_2017; @Robinson_Garcia_2020]. 
However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscriptions spending and the administrative efforts required to handle payments [@Pinfield_2016].
Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and publication fees, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020].
These transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles [@Hinchliffe_2019]. 
At the same time, they aim at a higher degree of transparency than previous subscription-only Big Deals, which bundled access to large journal portfolios of commercial publishers [@Bergstrom_2014].
However, the focus of transformative agreements remains primarily on these publishers with large hybrid journals and subscription-only portfolios [@Borrego_2020]. 
<!-- national negotiations, oa2020 -->
For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].

The introduction of transformative agreements aligns with funding policy changes, such as  the decision made by the  cOAlition S, a consortium of national funders including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
 According ot its Plan S, they only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018]. 
Specifically, cOAlition S funders agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
The German Research Foundation (DFG) has also extended its financial support for hybrid open access through transformative agreements, despite not being part of cOAlition S. Previously, the DFG only provided funding for full open access journals [@Jahn_2016].

Advocates of these agreements argue that they not only overcome the issue standardiazation of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.  A few agreements can impact an institution's open access share, resulting in a significant proportion of open access content. increase transparency

However, opponents of transformative agreements perceive continued reliance on big deals as problematic. They argue that the focus on large publishers perpetuates the business relationships that led to the serial crisis and lacks evidence that this leads to a reduction of In addition, concerns over inequality abound. trnasparency differ and now nsns

could not benn mass removal

Effects can be seen, mixed consequences. some countries who have introduced this fundign effects netherlands, sweden. no in germany, elsevier deal to price reduction and first steps towards mroe privacy. however, main goals associated woth trnasofrtiomative, the transition of the journal landscape not achivned, publsiehrs were not able to meet growtht raget, sweden did not see the flipping and has

This study contributes

@@

>>>>

For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
The idea was that when journals increasingly publish open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access. 
However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
Other publishers also recorded a low uptake.
In 2011, only 1-2% of articles in over 4,300 hybrid journals from large publishers were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].

With the introduction of central funding mechanisms for hybrid open access publication fees in some European countries since 2012 an increase in open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
In particular, British (17%), Austrian (15%) and Dutch (13%) universities contributed to this trend. 
However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and publication fees, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020; @Hinchliffe_2019].
Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
Although transformative agreements mainly focus on large hybrid journals and subscription-only portfolios of commercial publishers,  they aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023]. -->

The introduction of transformative agreements aligns with funding policy changes, such as the decision made by the cOAlition S, a consortium of national funders including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
According ot its Plan S, they only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018]. 
Specifically, cOAlition S funders agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
Notably, the German Research Foundation (DFG), despite not being part of cOAlition S, has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021]. Previously, the DFG only provided funding for full open access journals [@Jahn_2016].

By the end of 2023, many transformative agreements were implemented, but the outcomes were mixed.
The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch]. 
National consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Hinchliffe_2019; @Parmhed_2023]. 
The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019].
However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, perpetuating the business relationships that led to the serial crisis. 
Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023].
A substantial transition of hybrid journals towards full OA could not be observed [@Momeni_2021; @Matthias_2019], and the focus on large commercial publishers has led to concerns regarding inequality, because transformative agreements continue to target institutions from high-income countries [@Klebel_2023; @Butler_2023]. 
Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].

The controversies surrounding hybrid open access and transformative agreements have led to varying practical conclusions. 
For instance, the Swedish BIBSAM consortium recommended to only support agreements for publishing in full open access journals.
Likewise, most cOAlition S funders will discontinue financial support for transformative agreements from 2024 onwards [@Liverpool_2023].
The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to their failure to meet self-defined open access growth targets. 
In contrast, the German DEAL consortium signed a cost-effective five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.

Against this background, the research objective of this study is to investigate the open access uptake of hybrid journals between 2018 and 2022, and the extent to which open access in hybrid journals can be attributed to transformative agreements.
Previous quantitative science studies were able to investigate transformative agreements relative to specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] and journal portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2022; @Momeni_2023].
Others have focused soley on APC list prices [@Zhang_2022; @Shu_2023], not accounting for the peculiarities of transformative agreements. 
Similarly, bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.

The present study aims at a large-scale investigation using openly available data sources to answer the following research questions:

- What are the number and proportion of open access articles in hybrid journals published under transformative agreements between 2018-2022?
- Who is publishing open access in hybrid journals enabled by transformative agreements, and who is not?

The results of the study will be group by publishers, subjects and countries to account for the controversies.

@@

[^esacmarketwatch]: <https://esac-initiative.org/market-watch/>

### 2024-02-14 analysis/manuscript.tex
continue introduction ii

For over two decades, hybrid open access journal publishing, which makes
some articles openly available while others remain behind a paywall, has
been discussed as a model for transitioning the subscription system to
full open access (Prosser, 2003). The idea was that when journals
publish more and more open access articles, they could reduce revenues
from subscriptions, while libraries and funders could change their
funding models and shift expenditures from subscription to open access.
However, initial approaches, mainly based on publication fees, also
called article processing charges (APCs), did not contribute
substantially to an large open access uptake. In 2009, Springer reported
open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
al., 2010). Other publishers also recorded a low uptake. In 2011, only
1-2\% of articles in over 4,300 hybrid journals from large publishers
were open access (Björk, 2012), growing to around 4\% between 2011 and
2013 (Laakso \& Björk, 2016).

With the introduction of central funding mechanisms for hybrid open
access publication fees in a few countries since 2012, notably in the
UK, an increase in open access could be observed (Björk, 2017; Jubb et
al., 2017; Robinson-Garcia et al., 2020). However, this shift in funding
policy towards hybrid open access also added to the overall cost of
publishing, which includes subscriptions spending and the administrative
efforts required to handle payments (Pinfield et al., 2016). Moreover,
established large commercial publishers, which already dominated the
publishing market (Larivière et al., 2015), disproportionately benefited
from hybrid open access funding in comparison to full open access
publishers (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
Larivière, 2023).

As a consequence, libraries and their consortia began to develop
licensing strategies aimed at avoiding such `double dipping' scenarios,
in which well-established commercial publishers gain twice from reading
and publication fees, as well as to increase publisher-provided
immediate open access (BJÖRK \& SOLOMON, 2014; Schimmer et al., 2015).
These considerations resulted into transformative agreements, which
cover a broad range of successfully negotiated contracts between library
consortia and publishers from mid-2010s onwards where institutional
spending for subscriptions are re-purposed for open access publishing
(Borrego et al., 2020). These transformative agreements seek to control
costs while allowing a transitional phase for publishing more open
access articles (Hinchliffe, 2019). At the same time, they aim at a
higher degree of transparency than previous subscription-only Big Deals,
which bundled access to large journal portfolios of commercial
publishers (Bergstrom et al., 2014). However, the focus of
transformative agreements remains primarily on these publishers with
large hybrid journals and subscription-only portfolios (Borrego et al.,
2020). For example, the nationwide DEAL consortium in Germany, which
includes universities and academic institutions, collectively cancelled
their subscriptions with Elsevier from 2017 onwards to negotiate a
transformative agreement (Fraser et al., 2023).

The introduction of transformative agreements aligns with funding policy
changes, such as the decision made by the cOAlition S, a consortium of
national funders including the European Commission, to no longer provide
financial support for individual publication fees when publishing in
hybrid journals. According ot its Plan S, they only accept hybrid open
access through transformative agreements ``during a transition period
that should be as short as possible'' (Schiltz, 2018). Specifically,
cOAlition S funders agreed to support hybrid open access solely through
transformative agreements from 1 January 2021, until the end of 2024.
The German Research Foundation (DFG) has also extended its financial
support for hybrid open access through transformative agreements,
despite not being part of cOAlition S. Previously, the DFG only provided
funding for full open access journals (Jahn \& Tullney, 2016).

Advocates of these agreements argue that they not only overcome the
issue standardiazation of double-dipping but also help to shift library
spending from subscriptions to publications, streamline expenditures,
and offer more value than big deals, which focus on large publishers. A
few agreements can impact an institution's open access share, resulting
in a significant proportion of open access content. increase
transparency

However, opponents of transformative agreements perceive continued
reliance on big deals as problematic. They argue that the focus on large
publishers perpetuates the business relationships that led to the serial
crisis and lacks evidence that this leads to a reduction of In addition,
concerns over inequality abound. trnasparency differ and now nsns

could not benn mass removal

Effects can be seen, mixed consequences. some countries who have
introduced this fundign effects netherlands, sweden. no in germany,
elsevier deal to price reduction and first steps towards mroe privacy.
however, main goals associated woth trnasofrtiomative, the transition of
the journal landscape not achivned, publsiehrs were not able to meet
growtht raget, sweden did not see the flipping and has

This study contributes

@@

Open access articles in hybrid journals were identified through Creative
Commons (CC) license information in Crossref metadata. License
information relative to the ``accepted manuscript (AM)'' version were
not considered. Crossref was used for open access identification because
transformative agreements workflows generally require publishers to
deliver CC license information to this DOI registration agency (Geschuhn
\& Stone, 2017). Comparing Crossref license coverage with OpenAlex,
which re-uses open access evidence from Unpaywall, a widely used open
access discovery service that also parses journal websites for open
content licenses (Piwowar et al., 2018), highlighted ongoing challenges
to identify hybrid open access (Butler et al., 2023; Jahn et al., 2021;
Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
with CC license were retrieved using Crossref, while 950,260 articles
were tagged as ``hybrid'' according to the OpenAlex November 2023
release, which was used throughout this study. The biggest differences
concerned articles published between 2018 and 2020. In 2022, however,
Crossref and OpenAlex open access numbers only differ slightly (249,511
records using Crossref vs.~255,344 in OpenAlex). Notable difference
could be furthermore observed among some publishers that presumably did
not provide CC license information to Crossref including AIP Publishing,
American Physiological Society and Emerald. Crossref license metadata
was more complete with regard to articles from the publisher Wiley and
American Chemical Society. Finally, inconsistent open access status
information in previous OpenAlex versions was observed (Jahn et al.,
2023). According to the OpenAlex release notes, fixing this issue was
still ongoing, which might also explain this discrepancy.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bergstrom_2014}{}}%
Bergstrom, T. C., Courant, P. N., McAfee, R. P., \& Williams, M. A.
(2014). Evaluating big deal journal bundles. \emph{Proceedings of the
National Academy of Sciences}, \emph{111}(26), 9425--9430.
\url{https://doi.org/10.1073/pnas.1403006111}

@@

\leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
BJÖRK, B., \& SOLOMON, D. (2014). How research funders can finance APCs
in full OA and hybrid journals. \emph{Learned Publishing}, \emph{27}(2),
93--103. \url{https://doi.org/10.1087/20140203}

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
Jahn, N., Matthias, L., \& Laakso, M. (2021). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of elsevier. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

@@

@@

@@

@@

@@

@@

>>>>

For over two decades, hybrid open access journal publishing, which makes
some articles openly available while others remain behind a paywall, has
been discussed as a model for transitioning the subscription system to
full open access (Prosser, 2003). The idea was that when journals
increasingly publish open access articles, they could reduce revenues
from subscriptions, while libraries and funders could change their
funding models and shift expenditures from subscription to open access.
However, initial approaches, mainly based on publication fees, also
called article processing charges (APCs), did not contribute
substantially to an large open access uptake. In 2009, Springer reported
open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
al., 2010). Other publishers also recorded a low uptake. In 2011, only
1-2\% of articles in over 4,300 hybrid journals from large publishers
were open access (B.-C. Björk, 2012), growing to around 4\% between 2011
and 2013 (Laakso \& Björk, 2016).

With the introduction of central funding mechanisms for hybrid open
access publication fees in some European countries since 2012 an
increase in open access could be observed (B.-C. Björk, 2017; Huang et
al., 2020; Jubb et al., 2017; Piwowar et al., 2018). For example,
studying university output, Robinson-Garcia et al. (2020) estimated a
median uptake of 7.1\% in the period 2014-2017. In particular, British
(17\%), Austrian (15\%) and Dutch (13\%) universities contributed to
this trend. However, this shift in funding policy towards hybrid open
access also added to the overall cost of publishing, which includes
subscription spending and the administrative efforts required to handle
payments (Pinfield et al., 2016). Moreover, established large commercial
publishers, which already dominated the publishing market (Larivière et
al., 2015), disproportionately benefited from hybrid open access funding
in comparison to full open access publishers (Butler et al., 2023; Jahn
\& Tullney, 2016; Shu \& Larivière, 2023).

As a consequence, libraries and their consortia began to develop
licensing strategies aimed at avoiding such `double dipping' scenarios,
in which well-established commercial publishers gain twice from reading
and publication fees, as well as to increase publisher-provided
immediate open access (B. Björk \& Solomon, 2014; Schimmer et al.,
2015). These considerations resulted into transformative agreements,
which cover a broad range of successfully negotiated contracts between
library consortia and publishers from mid-2010s onwards where
institutional spending for subscriptions are re-purposed for open access
publishing (Borrego et al., 2020; Hinchliffe, 2019). Transformative
agreements seek to control costs while allowing a transitional phase for
publishing more open access articles. Although transformative agreements
mainly focus on large hybrid journals and subscription-only portfolios
of commercial publishers, they aim at a higher degree of transparency
than previous big deals, where contracts including payments were
confidential (Bergstrom et al., 2014).

The introduction of transformative agreements aligns with funding policy
changes, such as the decision made by the cOAlition S, a consortium of
national funders including the European Commission, to no longer provide
financial support for individual publication fees when publishing in
hybrid journals. According ot its Plan S, they only accept hybrid open
access through transformative agreements ``during a transition period
that should be as short as possible'' (Schiltz, 2018). Specifically,
cOAlition S funders agreed to support hybrid open access solely through
transformative agreements from 1 January 2021, until the end of 2024.
Notably, the German Research Foundation (DFG), despite not being part of
cOAlition S, has also extended its financial support for hybrid open
access through transformative agreements (Mittermaier, 2021).
Previously, the DFG only provided funding for full open access journals
(Jahn \& Tullney, 2016).

By the end of 2023, many transformative agreements were implemented, but
the outcomes are mixed. The ESAC Transformative Agreement
Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
the largest source of disclosure, recorded more than 800 transformative
agreements, resulting in up to 900.000 open access articles published in
both full open access and hybrid journals according to the accompanying
ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
National consortia reported increased open access volume, streamlined
payment and monitoring procedures, as well as extensive utilization of
open access options by the researchers they serve (Hinchliffe, 2019;
Marques \& Stone, 2020; Parmhed \& Säll, 2023). The ongoing
standardisation of transformative agreements contributed to improved
transparency in terms of contracts and publisher-provided article
metadata (Marques et al., 2019). However, with the growing trend toward
transformative agreements, continued reliance on big deals is perceived
as problematic, perpetuating the business relationships that led to the
serial crisis. Whether transformative agreements lead to reduced pricing
remains uncertain (\textbf{Borrego\_2023?}). A substantial transition of
hybrid journals towards full OA could not be observed (Matthias et al.,
2019; Momeni et al., 2021), and the focus on large commercial publishers
has led to concerns regarding inequality, because transformative
agreements continue to target institutions from high-income countries
(Butler et al., 2023; Klebel \& Ross-Hellauer, 2023). Besides, an
editorial-board resignation raised concerns that publishers' desire to
maximize journal publication volume ``without regard to quality'' is a
consequence of transformative agreements (Rasmussen, 2023).

The controversies surrounding hybrid open access and transformative
agreements have led to varying practical conclusions. For instance, the
Swedish BIBSAM consortium recommended to only support agreements for
publishing in full open access journals. Likewise, most cOAlition S
funders will discontinue financial support for transformative agreements
from 2024 onwards (Liverpool, 2023). The consortium also removed most
hybrid journals from its Transformative Journal program in 2023 due to
their failure to meet self-defined open access growth targets. In
contrast, the German DEAL consortium signed a cost-effective five-year
transformative agreement with Elsevier starting in 2024, while also
renewing its contracts with Springer Nature and Wiley until the end of
2028.

Against this background, the research objective of this study is to
investigate the open access uptake of hybrid journals between 2018 and
2022, and the extent to which open access in hybrid journals can be
attributed to transformative agreements. Previous quantitative science
studies were able to investigate transformative agreements relative to
specific countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et
al., 2020; Taubert et al., 2023; Wenaas, 2022) and journal portfolios
(Bakker et al., 2024; Fraser et al., 2023; Momeni et al., 2023; Pieper
\& Broschinski, 2018; \textbf{Jahn\_2022?}). Others have focused soley
on APC list prices (Shu \& Larivière, 2023; Zhang et al., 2022), not
accounting for the peculiarities of transformative agreements.
Similarly, bibliometric databases, even though many allow the retrieval
of open access articles in hybrid journals, do not directly attribute
them to specific transformative agreements.

The present study aims at a large-scale investigation using openly
available data sources to answer the following research questions:

\begin{itemize}
\tightlist
\item
  What are the number and proportion of open access articles in hybrid
  journals published under transformative agreements between 2018-2022?
\item
  Who is publishing open access in hybrid journals enabled by
  transformative agreements, and who is not?
\end{itemize}

The results of the study will be group by publishers, subjects and
countries to account for the controversies.

@@

Open access articles in hybrid journals were identified through Creative
Commons (CC) license information in Crossref metadata. License
information relative to the ``accepted manuscript (AM)'' version were
not considered. Crossref was used for open access identification because
transformative agreements workflows generally require publishers to
deliver CC license information to this DOI registration agency (Geschuhn
\& Stone, 2017). Comparing Crossref license coverage with OpenAlex,
which re-uses open access evidence from Unpaywall, a widely used open
access discovery service that also parses journal websites for open
content licenses (Piwowar et al., 2018), highlighted ongoing challenges
to identify hybrid open access (Butler et al., 2023; Jahn et al., 2022;
Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
with CC license were retrieved using Crossref, while 950,260 articles
were tagged as ``hybrid'' according to the OpenAlex November 2023
release, which was used throughout this study. The biggest differences
concerned articles published between 2018 and 2020. In 2022, however,
Crossref and OpenAlex open access numbers only differ slightly (249,511
records using Crossref vs.~255,344 in OpenAlex). Notable difference
could be furthermore observed among some publishers that presumably did
not provide CC license information to Crossref including AIP Publishing,
American Physiological Society and Emerald. Crossref license metadata
was more complete with regard to articles from the publisher Wiley and
American Chemical Society. Finally, inconsistent open access status
information in previous OpenAlex versions was observed (Jahn et al.,
2023). According to the OpenAlex release notes, fixing this issue was
still ongoing, which might also explain this discrepancy.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bakker_2024}{}}%
Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
transformative agreements on publication patterns: An analysis based on
agreements from the ESAC registry. \emph{International Journal of
Librarianship}, \emph{8}(4), 67--96.
\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}

\leavevmode\vadjust pre{\hypertarget{ref-Bergstrom_2014}{}}%
Bergstrom, T. C., Courant, P. N., McAfee, R. P., \& Williams, M. A.
(2014). Evaluating big deal journal bundles. \emph{Proceedings of the
National Academy of Sciences}, \emph{111}(26), 9425--9430.
\url{https://doi.org/10.1073/pnas.1403006111}

@@

\leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
Björk, B., \& Solomon, D. (2014). How research funders can finance APCs
in full OA and hybrid journals. \emph{Learned Publishing}, \emph{27}(2),
93--103. \url{https://doi.org/10.1087/20140203}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Haucap_2021}{}}%
Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
german {``DEAL''} on competition in the academic publishing market.
\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
\url{https://doi.org/10.1002/mde.3493}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Huang_2020}{}}%
Huang, C.-K. (Karl), Neylon, C., Hosking, R., Montgomery, L., Wilson, K.
S., Ozaygen, A., \& Brookes-Kenworthy, C. (2020). Evaluating the impact
of open access policies on research institutions. \emph{{eLife}},
\emph{9}. \url{https://doi.org/10.7554/elife.57067}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of elsevier. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Klebel_2023}{}}%
Klebel, T., \& Ross-Hellauer, T. (2023). The APC-barrier and its effect
on stratification in open access publishing. \emph{Quantitative Science
Studies}, \emph{4}(1), 22--43. \url{https://doi.org/10.1162/qss_a_00245}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Liverpool_2023}{}}%
Liverpool, L. (2023). Open-access reformers launch next bold publishing
plan. \emph{Nature}, \emph{623}(7986), 238--240.
\url{https://doi.org/10.1038/d41586-023-03342-6}

\leavevmode\vadjust pre{\hypertarget{ref-Marques_2020}{}}%
Marques, M., \& Stone, G. (2020). Transitioning to open access: An
evaluation of the {UK} springer compact agreement pilot 2016--2018.
\emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
\url{https://doi.org/10.5860/crl.81.6.913}

\leavevmode\vadjust pre{\hypertarget{ref-Marques_2019}{}}%
Marques, M., Woutersen-Windhouwer, S., \& Tuuliniemi, A. (2019).
Monitoring agreements with open access elements: Why article-level
metadata are important. \emph{Insights the {UKSG} Journal}, \emph{32}.
\url{https://doi.org/10.1629/uksg.489}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Matthias_2019}{}}%
Matthias, L., Jahn, N., \& Laakso, M. (2019). The two-way street of open
access journal publishing: Flip it and reverse it. \emph{Publications},
\emph{7}(2), 23. \url{https://doi.org/10.3390/publications7020023}

\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2021}{}}%
Mittermaier, B. (2021). {Rolle des Open Access Monitor Deutschland bei
der Antragstellung im DFG-Förderprogramm
Open-Access-Publikationskosten}. \emph{{O-Bib. Das Offene
Bibliotheksjournal / Herausgeber VDB}}, \emph{8}.
\url{https://doi.org/10.5282/O-BIB/5731}

\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2023}{}}%
Momeni, F., Dietze, S., Mayr, P., Biesenbender, K., \& Peters, I.
(2023). Which factors are associated with open access publishing? A
springer nature case study. \emph{Quantitative Science Studies},
\emph{4}(2), 353--371. \url{https://doi.org/10.1162/qss_a_00253}

\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2021}{}}%
Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
when a journal converts to open access? A bibliometric analysis.
\emph{Scientometrics}, \emph{126}(12), 9811--9827.
\url{https://doi.org/10.1007/s11192-021-03972-5}

\leavevmode\vadjust pre{\hypertarget{ref-Parmhed_2023}{}}%
Parmhed, S., \& Säll, J. (2023). Transformative agreements and their
practical impact: A librarian perspective. \emph{Insights the UKSG
Journal}, \emph{36}. \url{https://doi.org/10.1629/uksg.612}

\leavevmode\vadjust pre{\hypertarget{ref-Pieper_2018}{}}%
Pieper, D., \& Broschinski, C. (2018). {OpenAPC}: A contribution to a
transparent and reproducible monitoring of fee-based open access
publishing across institutions and nations. \emph{Insights the {UKSG}
Journal}, \emph{31}. \url{https://doi.org/10.1629/uksg.439}

@@

\leavevmode\vadjust pre{\hypertarget{ref-P_l_nen_2020}{}}%
Pölönen, J., Laakso, M., Guns, R., Kulczycki, E., \& Sivertsen, G.
(2020). Open access at the national level: A comprehensive analysis of
publications by finnish researchers. \emph{Quantitative Science
Studies}, \emph{1}(4), 1396--1428.
\url{https://doi.org/10.1162/qss_a_00084}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
Rasmussen, K. B. (2023). Interview with robert {``bob''} e. goodin.
\emph{Tidskrift För Politisk Filosofi}.
\url{https://www.politiskfilosofi.se/fulltext/2023-2/pdf/TPF_2023-2_interview_with_robert_bob_e_goodin.pdf}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Taubert_2023}{}}%
Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
Understanding differences of the OA uptake within the german university
landscape (2010--2020): Part 1---journal-based OA.
\emph{Scientometrics}, \emph{128}(6), 3601--3625.
\url{https://doi.org/10.1007/s11192-023-04716-3}

\leavevmode\vadjust pre{\hypertarget{ref-Wenaas_2022}{}}%
Wenaas, L. (2022). Choices of immediate open access and the relationship
to journal ranking and publish-and-read deals. \emph{Frontiers in
Research Metrics and Analytics}, \emph{7}.
\url{https://doi.org/10.3389/frma.2022.943932}
